Make any stupid plot you want

### 2021-02-17 draft/sections/intro.tex
invariance, quality metrics

The visual properties were first codified by Bertin\cite{bertinSemiologyGraphicsDiagrams2011a}, who also described the types of data that paired with the visual properties shown in figure~\ref{fig:intro_retinal_variables}. It is at this encoding level that Mackinlay formalized expressiveness as a homomorphic mapping which preserves some binary operator from one domain to another \cite{mackinlayAUTOMATICDESIGNGRAPHICAL1987}. These visual encodings are composited into the point, line, and area graphical marks, as shown in figure~\ref{fig:intro_retinal_variables} correspond. Marks can be generalized to glyphs, which are graphical objects that convey one or more attributes of the data entity mapped to it\cite{ware2019information, munznerVisualizationAnalysisDesign2014}. As retinal variables and marks are the building blocks of most visualizations, they must be fully expressible in a framework for building visualization tools.

\subsection{contribution}
This work presents a mathematical model of the transformation from data to graphic representation and a proof of concept implementation. Specifically, this work contributes

\begin{enumerate}
    \item a functional oriented visualization tool architecture
    \item topology-preserving maps from data to graphic 
    \item monoidal action equivariant maps from component to visual variable 
    \item algebraic sum such that more complex visualizations can be built from simple ones 
    \item prototype built on Matplotlib's existing infrastructure for the rendering 
\end{enumerate}
\end{document}

>>>>

The visual properties were first codified by Bertin\cite{bertinSemiologyGraphicsDiagrams2011a}, who also described the types of data that paired with the visual properties shown in figure~\ref{fig:intro_retinal_variables}. It is at this encoding level that Mackinlay formalized expressiveness as a homomorphic mapping which preserves some binary operator from one domain to another \cite{mackinlayAUTOMATICDESIGNGRAPHICAL1987}. These visual encodings are composited into the point, line, and area graphical marks, as shown in figure~\ref{fig:intro_retinal_variables} correspond. Marks can be generalized to glyphs, which are graphical objects that convey one or more attributes of the data entity mapped to it\cite{ware2019information, munznerVisualizationAnalysisDesign2014}. As retinal variables and marks are the building blocks of most visualizations, they must be fully expressible in a framework for building visualization tools. By building visualization concepts into the core architecture, developers can incorporate assessments of the visualization,  such as as quality metrics\cite{bertiniQualityMetricsHighdimensional2011a} or invariance \cite{kindlmann2014algebraic} of visualizations with respect to graphical encoding choices.

\subsection{Contribution}
\note{should I mention the categorical framework here too?}
This work presents a mathematical model of the transformation from data to graphic representation and a proof of concept implementation. Specifically, this work contributes

\begin{enumerate}
    \item a functional oriented visualization tool architecture
    \item topology-preserving maps from data to graphic 
    \item monoidal action equivariant maps from component to visual variable 
    \item algebraic sum such that more complex visualizations can be built from simple ones 
    \item prototype built on Matplotlib's infrastructure
\end{enumerate}
In contrast to the algebraic basis for visualization design proposed by Kindlmann and Scheideggar\cite{kindlmann2014algebraic} or Mackinlay's APT framework, we are proposing a topological framework for building tools to build visualizations. We defer judgement of expressivity and effectiveness to developers building domain specific tools, but provide them the framework to do so. 
\end{document}
